### 2017-07-03 reasoning.tex
EDits.

A characteristic of RG reasoning is that stability of an assertion is
always proven w.r.t to $R$, and not $R^{*}$, although interference
may include multiple environment steps, and $R$ only captures a single
step. This is nonetheless sound due to the the inductive reasoning: if
$Q$ is preserved by every step of $R$, then $Q$ is preserved by
$R^{*}$, and vice-versa.  However, such reasoning does not
extend naturally to isolation-constrained interference ($\R$)
because $R^{*}$ modulo $\I$ is not same as $\R^{*}$; the former is a
transitive relation constrained by $\I$, whereas the latter is the
transitive closure of a $\I$-constrained relation. We therefore
introduce a side-condition on $\I$ that restores the equality. The
condition requires $\I$ to allow an intereference $R^{*}(\stg,\stg'')$,
for two database states $\stg$ and $\stg''$, only if it also allows
interference for every prefix of $R^{*}(\stg,\stg'')$. In other words,
if $\I$ disallows interference from $\stg$ to $\stg'$, then an
$R$-step from $\stg'$ to $\stg''$ shouldn't make the interference from
$\stg$ to $\stg''$ valid. We call this the stability condition on
$\I$, defined as below:
\begin{smathpar}
\begin{array}{lcl}
  \stable(R,\I) & \Leftrightarrow & \forall \stl,\stg,\stg',\stg''.~
  \neg\I(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I(\stl,\stg,\stg'')
\end{array}
\end{smathpar}
It can be easily verified that the above stability condition is
satisfied by the isolation axioms from Sec.~\ref{sec:isolation}. For
instance, $\I_{ss}$, the snapshot axiom, is stable because if
$\I_{ss}$ is invalid, then an interference has already modified a
record, and no further interference will restore the original record,
because original record bears the id of a transaction that has long
committed. Thus, $\I_{ss}$ remains invalid.

>>>>

A characteristic of RG reasoning is that stability of an assertion is
always proven w.r.t to $R$, and not $R^{*}$, although interference may
include multiple environment steps, and $R$ only captures a single
step. This is nonetheless sound due to the the inductive reasoning: if
$Q$ is preserved by every step of $R$, then $Q$ is preserved by
$R^{*}$, and vice-versa.  However, such reasoning does not extend
naturally to isolation-constrained interference because $R^{*}$ modulo
$\I$ is not same as $\R^{*}$; the former is a transitive relation
constrained by $\I$, whereas the latter is the transitive closure of a
$\I$-constrained relation. We therefore introduce a side-condition on
$\I$ that restores the equality. The condition requires $\I$ to allow
an intereference $R^{*}(\stg,\stg'')$, for two database states $\stg$
and $\stg''$, only if it also allows interference for every prefix of
$R^{*}(\stg,\stg'')$. In other words, if $\I$ disallows interference
from $\stg$ to $\stg'$, then an $R$-step from $\stg'$ to $\stg''$
shouldn't make the interference from $\stg$ to $\stg''$ valid. We call
this the stability condition on $\I$, defined as below:
\begin{smathpar}
\begin{array}{lcl}
  \stable(R,\I) & \Leftrightarrow & \forall \stl,\stg,\stg',\stg''.~
  \neg\I(\stl,\stg,\stg') \conj R(\stg',\stg'') \Rightarrow
  \neg\I(\stl,\stg,\stg'')
\end{array}
\end{smathpar}
It can be easily verified that the above stability condition is
satisfied by the isolation axioms from Sec.~\ref{sec:isolation}. For
instance, $\I_{ss}$, the snapshot axiom, is stable because if
$\I_{ss}$ is invalid, then an interference has already modified a
record, and no further interference will restore the original record,
because original record bears the id of a transaction that has long
committed. Thus, $\I_{ss}$ remains invalid.
